Persistence in functional topology and a correction to a theorem of Morse

### 2021-07-29 abstract.tex

During the 1930s, Marston Morse developed a vast generalization of what is commonly known as Morse theory relating the critical points of a semi-continuous functional with the topology of its sublevel sets.
Morse and Tompkins applied this body of work, referred to as functional topology, to prove the Unstable Minimal Surface Theorem in the setting defined by Douglas' solution to Plateau's Problem.
Several concepts introduced by Morse in this context can be seen as early precursors to the theory of persistent homology, which by now has established itself as a popular tool in applied and theoretical mathematics.
In this article, we provide a modern redevelopment of the homological aspects of Morse's functional topology from the perspective of persistence theory.
We adjust several key definitions and prove stronger statements, including a generalized version of the Morse inequalities, in order to allow for novel uses of persistence techniques in functional analysis and symplectic geometry.
As an application, we identify and correct a mistake in the proof of the Unstable Minimal Surface Theorem by Morse and Tompkins.

>>>>

\begin{abstract}
	During the 1930s, Marston Morse developed a vast generalization of what is commonly known as Morse theory relating the critical points of a semi-continuous functional with the topology of its sublevel sets.
	Morse and Tompkins applied this body of work, referred to as functional topology, to prove the Unstable Minimal Surface Theorem in the setting defined by Douglas' solution to Plateau's Problem.
	Several concepts introduced by Morse in this context can be seen as early precursors to the theory of persistent homology, which by now has established itself as a popular tool in applied and theoretical mathematics.
	In this article, we provide a modern redevelopment of the homological aspects of Morse's functional topology from the perspective of persistence theory.
	We adjust several key definitions and prove stronger statements, including a generalized version of the Morse inequalities, in order to allow for novel uses of persistence techniques in functional analysis and symplectic geometry.
	As an application, we identify and correct a mistake in the proof of the Unstable Minimal Surface Theorem by Morse and Tompkins.
\end{abstract}

### 2021-07-29 connectivity.tex

% \section{Sufficient conditions for the existence of persistence diagrams} \label{s:connectivity}

@@

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is called \emph{%strongly
	locally homologically small} or \emph{%strongly
	$\HLC$} if for any $x \in X$, any neighborhood $V$ of $x$, and any pair of indices $s,t$ with $f(x) < s < t$ there is a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}

@@

Clearly, any $\HLC$ sublevel set filtration is also weakly $\HLC$:
while the weak $\HLC$ property merely requires the existence of an index $s \in (f(x),t)$ satisfying the $\HS$ condition, the %\emph{strong} 
$\HLC$ property requires the $\HS$ condition to hold for any $s \in (f(x),t)$.
If the filtration is induced by a continuous function, the converse also holds as the following theorem shows.

\begin{thm} \label{t:weak hlc to hlc}
	If the sublevel set filtration of a continuous function $f \colon X \to \R$ is weakly $\HLC$, then it is also %strongly 
	$\HLC$.
\end{thm}

@@

\begin{defi}
	For $u \in \R$, the sublevel set filtration of a function $f \colon X \to \R$ is called \emph{%strongly
	$\HLC$ below $u$} if for any $x \in X$, any neighborhood $V$ of $x$, and any pair of indices $s,t$ with $f(x) < s < t < u$ there is a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}

\begin{lem} \label{l:restriction of LHS filtration}
	Let $f \colon X \to \R$ be a function whose sublevel set filtration is %strongly 
	$\HLC$.
	Fix $u \in \R$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y = f_{\leq u}$.
	Then the sublevel set filtration defined by $g$ is %strongly 
	$\HLC$ below $u$.
\end{lem}

@@

	Since $Y \subseteq X$ carries the subspace topology, we may choose a neighborhood $V'$ of $x$ in $X$ such that $V = V' \cap Y$. 
	The sublevel set filtration of $f$ is assumed to be %strongly 
	$\HLC$, so there is a neighborhood $U' \subseteq V'$ of $x$ in $X$ such that the inclusion $f_{\leq s} \cap U' \to f_{\leq t} \cap V'$ is $\HS$.
	We set $U = U' \cap Y$, which defines a neighborhood of $x$ in $Y$.

@@

\begin{lem} \label{l:key lemma for q-tameness}
	Let $f \colon X \to \R$ be a function on a locally compact space $X$ whose sublevel set filtration is compact and %strongly 
	$\HLC$ below $u \in \R$, and consider subsets $C \subseteq L \subseteq X$ with $C$ compact and $L$ open.
	For any $s < t < u$ the inclusion $C \cap f_{\leq s} \to L \cap f_{\leq t}$ is $\HS$.
\end{lem}

@@

	Next, we will show using the %strong 
	$\HLC$ property that $\Sigma_{s, t}$ has the first required property, i.e., that any point $x \in L \cap f_{\leq s}$ has a neighborhood in $\Sigma_{s, t}$.
	Choose an arbitrary $s' \in (s,\, t)$.
	Since the sublevel set filtration of $f$ is %strongly
	$\HLC$ below $u$ and we have $s' < t < u$, there is an open neighborhood $U_x \subseteq L$ such that the inclusion
	$U_x \cap f_{\leq s'} \to L \cap f_{\leq t}$
	is $\HS$, so in particular $\HS_n$.
	By local compactness of $X$ we can choose a compact neighborhood $K_x$ of $x$ contained in $U_x$.
	Now $V_x = \interior (K_x)$ is a neighborhood of $x$ with $V_x \in \Sigma_{s,t}$.

@@

	In the open case, notice that for both choices of $i$ the inclusions
	$U_i \cap f_{\leq s''} \to L \cap f_{\leq t}$
	are $\HS_n$.
	Additionally, the inclusion
	$V'_1 \cap V'_2 \cap f_{\leq s'} \to U_1 \cap U_2 \cap f_{\leq s''}$
	is $\HS_{n-1}$ because it factors through the inclusion
	$K'_1 \cap K'_2 \cap f_{\leq s'} \to U_1 \cap U_2 \cap f_{\leq s''}$,
	which is $\HS_{n-1}$ by the induction hypothesis.
	Because the $V_i$ and $V'_i$ are open and because $\H$ has the open Mayer-Vietoris property, we obtain the following commutative diagram satisfying the assumptions of \cref{l:commutative algebra}:
	\begin{equation*}
	\begin{tikzcd}[column sep=small]
	\H_n(L \cap f_{\leq t}) \oplus \H_n(L \cap f_{\leq t}) \arrow[r] &
	\H_n(L \cap f_{\leq t}) & \\
	\H_{n}(U_1 \cap f_{\leq s''}) \oplus \H_n(U_2 \cap f_{\leq s''}) \arrow[r] \arrow[u] & 
	\H_{n}((U_1 \cup U_2) \cap f_{\leq s''}) \arrow[r] \arrow[u] &
	\H_{n-1}(U_1 \cap U_2 \cap f_{\leq s''}) \\ & 
	\H_{n}((V'_1 \cup V'_2) \cap f_{\leq s'}) \arrow[r] \arrow[u] &
	\H_{n-1}(V'_1 \cap V'_2 \cap f_{\leq s'}). \arrow[u]
	\end{tikzcd}
	\end{equation*}
	We conclude that the inclusion 
	$\left(V'_1 \cup V'_2 \right) \cap f_{\leq s'} \to L \cap f_{\leq t}$ 
	is $\HS_n$, which finishes this part of the proof.

	In the compact case, we apply \cref{l:neighborhood third} again to obtain compact sets $K''_i$ such that
	\begin{equation*}
	V_i \subseteq K_i \subseteq V'_i \subseteq K'_i \subseteq V''_i \subseteq K''_i \subseteq U_i \subseteq L
	\end{equation*}
	where $V''_i = \interior(K''_i)$.
	The rest of the proof is then analogous to the previous case:
	We have that for both choices of~$i$ the inclusions
	$K''_i \cap f_{\leq s''} \to L \cap f_{\leq t}$
	are $\HS_n$ because this is true for the corresponding inclusions with $K''_i$ replaced by $U_i$.
	Moreover, the inclusion
	$K'_1 \cap K'_2 \cap f_{\leq s'} \to K''_1 \cap K''_2 \cap f_{\leq s''}$
	is $\HS_{n-1}$ because it factors through the inclusion
	$K'_1 \cap K'_2 \cap f_{\leq s'} \to V''_1 \cap V''_2 \cap f_{\leq s''}$,
	which is $\HS_{n-1}$ by the induction hypothesis.
	Because the $K'_i$ and $K''_i$ as well as the sublevel sets of $f$ are all compact and because $\H$ has the compact Mayer-Vietoris property, we obtain the following commutative diagram satisfying the assumptions of \cref{l:commutative algebra}:
	\begin{equation*}
	\begin{tikzcd}[column sep=small]
	\H_n(L \cap f_{\leq t}) \oplus \H_n(L \cap f_{\leq t}) \arrow[r] &
	\H_n(L \cap f_{\leq t}) & \\
	\H_{n}(K''_1 \cap f_{\leq s''}) \oplus \H_n(K''_2 \cap f_{\leq s''}) \arrow[r] \arrow[u] & 
	\H_{n}((K''_1 \cup K''_2) \cap f_{\leq s''}) \arrow[r] \arrow[u] &
	\H_{n-1}(K''_1 \cap K''_2 \cap f_{\leq s''}) \\ & 
	\H_{n}((K'_1 \cup K'_2) \cap f_{\leq s'}) \arrow[r] \arrow[u] &
	\H_{n-1}(K'_1 \cap K'_2 \cap f_{\leq s'}). \arrow[u]
	\end{tikzcd}
	\end{equation*}
	We conclude that the inclusion 
	$\left(K'_1 \cup K'_2 \right) \cap f_{\leq s'} \to L \cap f_{\leq t}$
	is $\HS_n$, so the same is true for the inclusion
	$\left(V'_1 \cup V'_2 \right) \cap f_{\leq s'} \to L \cap f_{\leq t}$ 
	because it factors through the previous one.
\end{proof}

We can now complete the proof of the claim stating that for compact sublevel set filtrations, %strong 
$\HLC$ implies q-tameness.

\begin{proof}[Proof of \cref{t:local connectedness implies q-tameness}]
	By definition, the sublevel set filtration of $f$ is q-tame if and only if the inclusion $f_{\leq s} \to f_{\leq t}$ is $\HS$ for all pairs $s < t$. 
	Choose $u \in \R$ with $u > t$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y=f_{\leq u}$. 
	Since we assume $f$ to induce a %strongly 
	$\HLC$ sublevel set filtration, by \cref{l:restriction of LHS filtration} the sublevel set filtration of $g$ is %strongly 
	$\HLC$ below $u$. 
	Clearly, the sublevel set filtration of $g$ is also compact, and its domain $Y$ is locally compact being a compact Hausdorff space by assumption.
	Thus, we can apply \cref{l:key lemma for q-tameness} to the filtration $g_{\leq \bullet}$ with $C = f_{\leq s}$ and $L = Y$ to obtain that the inclusion $f_{\leq s} = C \cap g_{\leq s} \to L \cap g_{\leq t} = f_{\leq t}$ is $\HS$.
\end{proof}

>>>>

@@

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is called \emph{locally homologically small} or \emph{$\HLC$} if for any $x \in X$, any neighborhood $V$ of $x$, and any pair of indices $s,t$ with $f(x) < s < t$ there is a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}

@@

Clearly, any $\HLC$ sublevel set filtration is also weakly $\HLC$:
while the weak $\HLC$ property merely requires the existence of an index $s \in (f(x),t)$ satisfying the $\HS$ condition, the $\HLC$ property requires the $\HS$ condition to hold for any $s \in (f(x),t)$.
If the filtration is induced by a continuous function, the converse also holds as the following theorem shows.

\begin{thm} \label{t:weak hlc to hlc}
	If the sublevel set filtration of a continuous function $f \colon X \to \R$ is weakly $\HLC$, then it is also $\HLC$.
\end{thm}

@@

\begin{defi}
	For $u \in \R$, the sublevel set filtration of a function $f \colon X \to \R$ is called \emph{$\HLC$ below $u$} if for any $x \in X$, any neighborhood $V$ of $x$, and any pair of indices $s,t$ with $f(x) < s < t < u$ there is a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}

\begin{lem} \label{l:restriction of LHS filtration}
	Let $f \colon X \to \R$ be a function whose sublevel set filtration is $\HLC$.
	Fix $u \in \R$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y = f_{\leq u}$.
	Then the sublevel set filtration defined by $g$ is $\HLC$ below $u$.
\end{lem}

@@

	Since $Y \subseteq X$ carries the subspace topology, we may choose a neighborhood $V'$ of $x$ in $X$ such that $V = V' \cap Y$.
	The sublevel set filtration of $f$ is assumed to be $\HLC$, so there is a neighborhood $U' \subseteq V'$ of $x$ in $X$ such that the inclusion $f_{\leq s} \cap U' \to f_{\leq t} \cap V'$ is $\HS$.
	We set $U = U' \cap Y$, which defines a neighborhood of $x$ in $Y$.

@@

\begin{lem} \label{l:key lemma for q-tameness}
	Let $f \colon X \to \R$ be a function on a locally compact space $X$ whose sublevel set filtration is compact and $\HLC$ below $u \in \R$, and consider subsets $C \subseteq L \subseteq X$ with $C$ compact and $L$ open.
	For any $s < t < u$ the inclusion $C \cap f_{\leq s} \to L \cap f_{\leq t}$ is $\HS$.
\end{lem}

@@

	Next, we will show using the $\HLC$ property that $\Sigma_{s, t}$ has the first required property, i.e., that any point $x \in L \cap f_{\leq s}$ has a neighborhood in $\Sigma_{s, t}$.
	Choose an arbitrary $s' \in (s,\, t)$.
	Since the sublevel set filtration of $f$ is $\HLC$ below $u$ and we have $s' < t < u$, there is an open neighborhood $U_x \subseteq L$ such that the inclusion
	$U_x \cap f_{\leq s'} \to L \cap f_{\leq t}$
	is $\HS$, so in particular $\HS_n$.
	By local compactness of $X$ we can choose a compact neighborhood $K_x$ of $x$ contained in $U_x$.
	Now $V_x = \interior (K_x)$ is a neighborhood of $x$ with $V_x \in \Sigma_{s,t}$.

@@

	In the open case, notice that for both choices of $i$ the inclusions
	$U_i \cap f_{\leq s''} \to L \cap f_{\leq t}$
	are $\HS_n$.
	Additionally, the inclusion
	$V'_1 \cap V'_2 \cap f_{\leq s'} \to U_1 \cap U_2 \cap f_{\leq s''}$
	is $\HS_{n-1}$ because it factors through the inclusion
	$K'_1 \cap K'_2 \cap f_{\leq s'} \to U_1 \cap U_2 \cap f_{\leq s''}$,
	which is $\HS_{n-1}$ by the induction hypothesis.
	Because the $V_i$ and $V'_i$ are open and because $\H$ has the open Mayer-Vietoris property, we obtain the following commutative diagram satisfying the assumptions of \cref{l:commutative algebra}:
	\begin{equation*}
	\begin{tikzcd}[column sep=small]
	\H_n(L \cap f_{\leq t}) \oplus \H_n(L \cap f_{\leq t}) \arrow[r] &
	\H_n(L \cap f_{\leq t}) & \\
	\H_{n}(U_1 \cap f_{\leq s''}) \oplus \H_n(U_2 \cap f_{\leq s''}) \arrow[r] \arrow[u] & 
	\H_{n}((U_1 \cup U_2) \cap f_{\leq s''}) \arrow[r] \arrow[u] &
	\H_{n-1}(U_1 \cap U_2 \cap f_{\leq s''}) \\ & 
	\H_{n}((V'_1 \cup V'_2) \cap f_{\leq s'}) \arrow[r] \arrow[u] &
	\H_{n-1}(V'_1 \cap V'_2 \cap f_{\leq s'}) \arrow[u].
	\end{tikzcd}
	\end{equation*}
	We conclude that the inclusion 
	$\left(V'_1 \cup V'_2 \right) \cap f_{\leq s'} \to L \cap f_{\leq t}$ 
	is $\HS_n$, which finishes this part of the proof.

	In the compact case, we apply \cref{l:neighborhood third} again to obtain compact sets $K''_i$ such that
	\begin{equation*}
	V_i \subseteq K_i \subseteq V'_i \subseteq K'_i \subseteq V''_i \subseteq K''_i \subseteq U_i \subseteq L
	\end{equation*}
	where $V''_i = \interior(K''_i)$.
	The rest of the proof is then analogous to the previous case:
	We have that for both choices of~$i$ the inclusions
	$K''_i \cap f_{\leq s''} \to L \cap f_{\leq t}$
	are $\HS_n$ because this is true for the corresponding inclusions with $K''_i$ replaced by $U_i$.
	Moreover, the inclusion
	$K'_1 \cap K'_2 \cap f_{\leq s'} \to K''_1 \cap K''_2 \cap f_{\leq s''}$
	is $\HS_{n-1}$ because it factors through the inclusion
	$K'_1 \cap K'_2 \cap f_{\leq s'} \to V''_1 \cap V''_2 \cap f_{\leq s''}$,
	which is $\HS_{n-1}$ by the induction hypothesis.
	Because the $K'_i$ and $K''_i$ as well as the sublevel sets of $f$ are all compact and because $\H$ has the compact Mayer-Vietoris property, we obtain the following commutative diagram satisfying the assumptions of \cref{l:commutative algebra}:
	\begin{equation*}
	\begin{tikzcd}[column sep=small]
	\H_n(L \cap f_{\leq t}) \oplus \H_n(L \cap f_{\leq t}) \arrow[r] &
	\H_n(L \cap f_{\leq t}) & \\
	\H_{n}(K''_1 \cap f_{\leq s''}) \oplus \H_n(K''_2 \cap f_{\leq s''}) \arrow[r] \arrow[u] & 
	\H_{n}((K''_1 \cup K''_2) \cap f_{\leq s''}) \arrow[r] \arrow[u] &
	\H_{n-1}(K''_1 \cap K''_2 \cap f_{\leq s''}) \\ & 
	\H_{n}((K'_1 \cup K'_2) \cap f_{\leq s'}) \arrow[r] \arrow[u] &
	\H_{n-1}(K'_1 \cap K'_2 \cap f_{\leq s'}) \arrow[u].
	\end{tikzcd}
	\end{equation*}
	We conclude that the inclusion 
	$\left(K'_1 \cup K'_2 \right) \cap f_{\leq s'} \to L \cap f_{\leq t}$
	is $\HS_n$, so the same is true for the inclusion
	$\left(V'_1 \cup V'_2 \right) \cap f_{\leq s'} \to L \cap f_{\leq t}$ 
	because it factors through the previous one.
\end{proof}

We can now complete the proof of the claim stating that for compact sublevel set filtrations, $\HLC$ implies q-tameness.

\begin{proof}[Proof of \cref{t:local connectedness implies q-tameness}]
	By definition, the sublevel set filtration of $f$ is q-tame if and only if the inclusion $f_{\leq s} \to f_{\leq t}$ is $\HS$ for all pairs $s < t$.
	Choose $u \in \R$ with $u > t$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y=f_{\leq u}$.
	Since we assume $f$ to induce a $\HLC$ sublevel set filtration, by \cref{l:restriction of LHS filtration} the sublevel set filtration of $g$ is $\HLC$ below $u$.
	Clearly, the sublevel set filtration of $g$ is also compact, and its domain $Y$ is locally compact being a compact Hausdorff space by assumption.
	Thus, we can apply \cref{l:key lemma for q-tameness} to the filtration $g_{\leq \bullet}$ with $C = f_{\leq s}$ and $L = Y$ to obtain that the inclusion $f_{\leq s} = C \cap g_{\leq s} \to L \cap g_{\leq t} = f_{\leq t}$ is $\HS$.
\end{proof}

### 2021-07-29 inequalities.tex

Working entirely in the algebraic setting, we will now fix a graded q-tame persistence module $M$.
Of course, one may think of $M$ as the persistent homology of a q-tame filtration, but $M$ could for example also arise as the filtered Floer homology of some Hamiltonian on a symplectic manifold.
By \cref{t:q-tame modules have barcodes}, $M$ has a persistence diagram in every degree $d$ with multiplicity function denoted $\m_d \colon \multiplicityDomain \to \N$.
In analogy to Morse's definitions, we may define the $(d, t, \epsilon)$-\emph{cap number} of our graded q-tame persistence module $M$ in terms of its persistence diagram as
\begin{equation*}
m_{d}^{\epsilon}(t) =
\sum_{\substack{p \in \R \\ t - p > \epsilon}} \m_{d-1}(p, t) +
\sum_{\substack{q \in \R \\ q - t > \epsilon}} \m_d(t, q).
\end{equation*}
Note that finiteness of these quantities is ensured by the q-tameness of $M$ and the use of a non-zero $e$ bounding below the persistence of the considered features.
To see the necessity of this second condition, consider the q-tame persistence module given by the infinite product $\prod_{n \in \N_{> 0}} C([0,1/n))$.

Whenever the sums below are well-defined, we also consider the \emph{$(d,\epsilon)$-cap numbers} 
\begin{equation} \label{e:cap numbers}
m_{d}^{\epsilon} =
\sum_{t} m_{d}^{\epsilon}(t) \ =
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ q \neq \infty}} \m_{d-1}(p,q) \ +
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ p \neq -\infty}}\m_d(p,q).
\end{equation}
An important setting where all cap numbers are well-defined is when $M$ is the persistent homology of the sublevel set filtration of a bounded function.
A more general statement can be made using the following notion.

@@

\begin{thm} \label{t:cap numbers well defined}
	Let $N$ be a q-tame persistence module that is both initially and eventually constant.
	If $\n$ is the multiplicity function of the persistence diagram of $N$,
	then 
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > \epsilon } } \n (p,q) < \infty
	\end{equation*}
	for each $\epsilon > 0$.
\end{thm}

\begin{proof}
    Let $t_0, t_1 \in \R$ be as in \cref{d:initially and eventually constant}.
	We split the sum whose finiteness we want to show in two parts
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > \epsilon } } \n (p,q)
	\ = \!\!
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > \epsilon } } \n (p,q)
	\ +
	\sum_{ \substack{ (p,q) \in T \\ q-p > \epsilon } } \n (p,q)
	\end{equation*}
	and show for each summand separately that it is finite.
	Here, $T$ denotes the triangle
	\begin{equation*}
	T = \{(p,q) \in \multiplicityDomain \mid t_0 \leq p < q \leq t_1\}.
	\end{equation*}

	For the first summand, observe that because $N$ is constant below $t_0$ and constant above $t_1$, we have $\n(p,q) = 0$ whenever one of $-\infty < p < t_0$ or $q < t_0$ or $p > t_1$ or $t_1 < q < \infty$ holds.
	This implies 
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > \epsilon } } \n (p,q)
	\ = \!
	\sum_{t_0 < q < t_1} \n (-\infty,q)
	\ + \!
	\sum_{t_0 < p < t_1} \n (p, \infty)
	\end{equation*}
	which is clearly finite because $N$ is q-tame.

@@

\begin{proof}
	Using a telescopic sum argument, the claimed inequalities can be seen to be equivalent to the existence of a sequence $(\nu_d)_d$ of non-negative integers with $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$.
	Since we assume $\m_d(-\infty, p) = 0$ for all $p$, we have
	\begin{align*}
	m^{\epsilon}_{d} \ &=
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ q \neq \infty}} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q) \ =
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon }} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q)
	\end{align*}
	and
	\begin{equation*}
	p_{d} \ = \!\!\!
	\sum_{p \in \R \cup \{-\infty\}} \m_d(p,\infty) \ = \
	\sum_{p \in \R} \m_d(p,\infty),
	\end{equation*}
	which yields
	\begin{align*}
	m^{\epsilon}_{d} - p_{d} \ &= 
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q) \ - \
	\sum_{p \in \R} \m_d(p,\infty) \ =
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_{d-1}(p,q)
	\ + \!
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_d(p,q),
	\end{align*}
	so the sequence given by $\nu_d = \sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_d(p,q)$ satisfies the equation $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$ as desired.
\end{proof}

>>>>

Working entirely in the algebraic setting, we will now fix a graded q-tame persistence module $M$.
Of course, one may think of $M$ as the persistent homology of a q-tame filtration, but $M$ could for example also arise as the filtered Floer homology of some Hamiltonian on a symplectic manifold.
By \cref{t:q-tame modules have barcodes}, $M$ has a persistence diagram in every degree $d$ with multiplicity function denoted $\m_d \colon \mulDom \to \N$.
In analogy to Morse's definitions, we may define the $(d, t, \epsilon)$-\emph{cap number} of our graded q-tame persistence module $M$ in terms of its persistence diagram as
\begin{equation*}
m_{d}^{\epsilon}(t) \ =
\sum_{\substack{p \in \R \\ t - p > \epsilon}} \m_{d-1}(p, t) +
\sum_{\substack{q \in \R \\ q - t > \epsilon}} \m_d(t, q).
\end{equation*}
Note that finiteness of these quantities is ensured by the q-tameness of $M$ and the use of a non-zero $e$ bounding below the persistence of the considered features.
To see the necessity of this second condition, consider the q-tame persistence module given by the infinite product $\prod_{n \in \N_{> 0}} C([0,1/n))$.

Whenever the sums below are well-defined, we also consider the \emph{$(d,\epsilon)$-cap numbers} 
\begin{equation} \label{e:cap numbers}
m_{d}^{\epsilon} =
\sum_{t} m_{d}^{\epsilon}(t) \ =
\sum_{\substack{(p,q) \in \mulDom \\ q - p > \epsilon \\ q \neq \infty}} \m_{d-1}(p,q) \ +
\sum_{\substack{(p,q) \in \mulDom \\ q - p > \epsilon \\ p \neq -\infty}}\m_d(p,q).
\end{equation}
An important setting where all cap numbers are well-defined is when $M$ is the persistent homology of the sublevel set filtration of a bounded function.
A more general statement can be made using the following notion.

@@

\begin{thm} \label{t:cap numbers well defined}
	Let $N$ be a q-tame persistence module that is both initially and eventually constant.
	If $\n$ is the multiplicity function of the persistence diagram of $N$,
	then 
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \mulDom \\ q-p > \epsilon } } \n (p,q) < \infty
	\end{equation*}
	for each $\epsilon > 0$.
\end{thm}

\begin{proof}
	Let $t_0, t_1 \in \R$ be as in \cref{d:initially and eventually constant}.
	We split the sum whose finiteness we want to show in two parts
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \mulDom \\ q-p > \epsilon } } \n (p,q)
	\ = \!\!
	\sum_{ \substack{ (p,q) \in \mulDom \setminus T \\ q-p > \epsilon } } \n (p,q)
	\ +
	\sum_{ \substack{ (p,q) \in T \\ q-p > \epsilon } } \n (p,q)
	\end{equation*}
	and show for each summand separately that it is finite.
	Here, $T$ denotes the triangle
	\begin{equation*}
	T = \{(p,q) \in \mulDom \mid t_0 \leq p < q \leq t_1\}.
	\end{equation*}

	For the first summand, observe that because $N$ is constant below $t_0$ and constant above $t_1$, we have $\n(p,q) = 0$ whenever one of $-\infty < p < t_0$ or $q < t_0$ or $p > t_1$ or $t_1 < q < \infty$ holds.
	This implies 
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \mulDom \setminus T \\ q-p > \epsilon } } \n (p,q)
	\ = \!
	\sum_{t_0 < q < t_1} \n (-\infty,q)
	\ + \!
	\sum_{t_0 < p < t_1} \n (p, \infty)
	\end{equation*}
	which is clearly finite because $N$ is q-tame.

@@

\begin{proof}
	Using a telescopic sum argument, the claimed inequalities can be seen to be equivalent to the existence of a sequence $(\nu_d)_d$ of non-negative integers with $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$.
	Since we assume $\m_d(-\infty, p) = 0$ for all $p$, we have
	\begin{align*}
	m^{\epsilon}_{d} \ &=
	\sum_{\substack{(p,q) \in \mulDom \\ q-p > \epsilon \\ q \neq \infty}} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \mulDom \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q) \ =
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon }} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \mulDom \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q)
	\end{align*}
	and
	\begin{equation*}
	p_{d} \ = \!\!\!
	\sum_{p \in \R \cup \{-\infty\}} \m_d(p,\infty) \ = \
	\sum_{p \in \R} \m_d(p,\infty),
	\end{equation*}
	which yields
	\begin{align*}
	m^{\epsilon}_{d} - p_{d} \ &= 
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \mulDom \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q) \ - \
	\sum_{p \in \R} \m_d(p,\infty) \ =
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_{d-1}(p,q)
	\ + \!
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_d(p,q),
	\end{align*}
	so the sequence given by $\nu_d = \sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_d(p,q)$ satisfies the equation $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$ as desired.
\end{proof}
